\chapter*{Acknowledgements}

I am very thankful to have had the opportunity of carried out this thesis.
This research was made possible by the great support of my supervisors and mentors Carmina G. Almudéver, Lingling Lao and Prof. Koen Bertels.
They accepted me besides my poor background on quantum computing and they were patient enough to teach me about it and about research.
I also would like to separately thank my supervisor in Barcelona, Antonio Rubio, who offered me and facilitated the opportunity of developing this study.

I appreciated all the help offered by my lab colleagues Hans van Someren, Imran Ashraf, Andreas Papageorgiou, Leon Riesebos and Aritra Sarkar.
We had interesting discussions that helped me to elucidate the solutions of the problems that appeared through out the research.
I also would like to thank Savvas Varsamopoulos, Abid Moueddene, Prof. Koen, Vieri Mattei, Miguel Serrão, Aritra, Leon and Andreas for all the fun and the beers we had together.

In general, I would like to express my gratitude to all the QCE department, specially to Erik de Vries -- who had deal with all my demands related with the lab servers --, Joyce van Velzen and Lidwina Tromp -- who arranged all the burocracy requirements in TU Delft for me.

Thanks for all your support!

--

Agradezco mucho la compañía de mi familia en este tiempo, quienes no han parado de cuidar de mí.
Sin ellos mi estancia en Los Países bajos hubiera sido inaguantable.
Doy gracias a mi madre y a mi hermana por todas esas videollamadas que me han hecho sentir tan cerca de casa.

Finalmente, quiero agradecer la compañía de todos mis amigos en éste tiempo.
Gracias por venir a visitarme y apoyarme también en la distancia.
Especialmente a Carlos Robisco, Cristina Puigjaner y Germán Fañanás, quienes me han ayudado mucho a aclararme con mi vocación.

¡Muchas gracias a todos por haberme acompañado en éste tiempo!

% \begin{comment}
\begin{flushright}
{\makeatletter\itshape
    \@author \\
    Delft, The Netherlands \\
    February 2019
\makeatother}
\end{flushright}
% \end{comment}
